HealpixNet: Efficient spherical Convolutional Neural Network with Healpix sampling for cosmological applications

### 2018-08-03 scnn.tex
some strange symbol
add architecture

    \item (TK) add small description of maps\_downsampled\_64.npz in the whole\_sphere notebook.
    \item (TK) update the README for data zip - file from NP
    \item (NP) Finalise readme on Github code
    \item (Waiting for Zenodo) (NP) Finalize the data on zenodo
    \item TK check part sphere notebook and whole sphere
    \item branch with notebooks with plots (results) NP
    \item NP move the repo to HealpixNet
    \item check if papers which appear as arxiv were not published in a conference or journal
    \item (TK) add info on comparison with histograms, so that it's just a demonstration
    \item TK read the result section
    \item TK check the appendix
    \item (almost done) TK optional - example of data (CMB (mask) + mass map + radio intensity)
    \item (almost done) TK read everything
    \item MD add irregular (random) sampling to comparison with spherical harmonics? Could motivate people to think about different samplings.
    \item uniformize text size in figures
    \item (done! first version done) conclusions
    \item (done!) Michael section 4.5 (efficient convolutions) and 3.2 (conv on graphs)
    \item (done!) (TK NP) if we have time make a new candidate for network plot (optional)
    \item (done!) TK read the heat section again
    \item (done!) (TK) make a new proposition for a plot Figure 6: Two levels of coarsening and pooling
    \item (TK took exec decision and moved it :D) Should the heat example go to appendix?
\end{itemize}

@@

From the HEALPix pixelization, we build a weighted undirected graph $\G = (\V, \E, \W)$, where $\V$ is the set of $N_{pix} = |\V|$ nodes, $\E$ is the set of edges, and $\W$ is the weighted adjacency matrix.
In our graph, each pixel $i$ is represented by a node (also called vertex) $v_i \in \V$.
Each node $v_i$ is then connected to the $8$ (or $7$)\footnote{\label{neighbors}The $12 \cdot 4 = 48$ pixels at the corner of each rhombus of the base dodecahedron only have 7 neighboring pixels. See Figures~\ref{fig:healpix_graph_4} and~\ref{fig:pooling}.} nodes $v_j$ which represent the neighboring pixels $j$ of pixel $i$, forming edges $(v_i, v_j) \in \E$. Given those edges, we define the weighted adjacency matrix $\W \in \R^{N_{pix} \times N_{pix}}$ as
\begin{equation*}
	\W_{ij} = \begin{cases}
		\exp \left( -\frac{\|\x_i-\x_j\|_2^2}{\sigma^2} \right) & \text{if pixels $i$ and $j$ are neighbors,} \\
		0 & \text{otherwise,} \\
	\end{cases}
\end{equation*}
where $\x_i$ is a vector encoding the 3-dimensional coordinates of pixel $i$, and
\begin{equation*}
	\sigma = \frac{1}{|\E|} \sum_{(v_i, v_j) \in \E} \|\x_i-\x_j\|_2
\end{equation*}
is the average Euclidean distance over all connected pixels. This weighting scheme is important as distances between pixels are not equal.
Other weighting schemes are possible. For example,~\cite{khasanova2017graphomni} uses the inverse of the distance instead. We found out that the one proposed above works well for our purpose, and did not investigate other approaches, leaving it to future work.
\figref{healpix_graph_4} shows a graph constructed from the HEALPix sampling of a sphere.

@@

\nati{4.8 is long... Should we go in such details?}
\todo{Notation GC32 - P256 - FC10 to describe an architecture? Maybe more useful in Experiments where we describe the architecture used for each experiment.}

@@

\subsection{Selected architecture}
\todo{Describe our fully convolutional architecture}

>>>>

    \item (TK) add small description of maps\_downsampled\_64.npz in the whole\_sphere notebook.
    \item (TK) update the README for data zip - file from NP
    \item (NP) Finalise readme on Github code
    \item TK check part sphere notebook and whole sphere
    \item branch with notebooks with plots (results) NP
    \item NP move the repo to HealpixNet
    \item check if papers which appear as arxiv were not published in a conference or journal
    \item (TK) add info on comparison with histograms, so that it's just a demonstration
    \item TK read the result section
    \item TK check the appendix
    \item (almost done) TK optional - example of data (CMB (mask) + mass map + radio intensity)
    \item (almost done) TK read everything
    \item MD add irregular (random) sampling to comparison with spherical harmonics? Could motivate people to think about different samplings.
    \item uniformize text size in figures
    \item (done! first version done) conclusions
    \item (done!) Michael section 4.5 (efficient convolutions) and 3.2 (conv on graphs)
    \item (done!) (TK NP) if we have time make a new candidate for network plot (optional)
    \item (done!) TK read the heat section again
    \item (done!) (TK) make a new proposition for a plot Figure 6: Two levels of coarsening and pooling
    \item (TK took exec decision and moved it :D) Should the heat example go to appendix?
\end{itemize}

@@

From the HEALPix pixelization, we build a weighted undirected graph $\G = (\V, \E, \W)$, where $\V$ is the set of $N_{pix} = |\V|$ nodes, $\E$ is the set of edges, and $\W$ is the weighted adjacency matrix.
In our graph, each pixel $i$ is represented by a node (also called vertex) $v_i \in \V$.
Each node $v_i$ is then connected to the $8$ (or $7$)\footnote{\label{neighbors}The $12 \cdot 4 = 48$ pixels at the corner of each rhombus of the base dodecahedron only have 7 neighboring pixels. See Figures~\ref{fig:healpix_graph_4} and~\ref{fig:pooling}.} nodes $v_j$ which represent the neighboring pixels $j$ of pixel $i$, forming edges $(v_i, v_j) \in \E$. Given those edges, we define the weighted adjacency matrix $\W \in \R^{N_{pix} \times N_{pix}}$ as
\begin{equation*}
	\W_{ij} = \begin{cases}
		\exp \left( -\frac{\|\x_i-\x_j\|_2^2}{\rho^2} \right) & \text{if pixels $i$ and $j$ are neighbors,} \\
		0 & \text{otherwise,} \\
	\end{cases}
\end{equation*}
where $\x_i$ is a vector encoding the 3-dimensional coordinates of pixel $i$, and
\begin{equation*}
	\rho = \frac{1}{|\E|} \sum_{(v_i, v_j) \in \E} \|\x_i-\x_j\|_2
\end{equation*}
is the average Euclidean distance over all connected pixels. This weighting scheme is important as distances between pixels are not equal.
Other weighting schemes are possible. For example,~\cite{khasanova2017graphomni} uses the inverse of the distance instead. We found out that the one proposed above works well for our purpose, and did not investigate other approaches, leaving it to future work.
\figref{healpix_graph_4} shows a graph constructed from the HEALPix sampling of a sphere.

@@

\todo{Notation GC32 - P256 - FC10 to describe an architecture? Maybe more useful in Experiments where we describe the architecture used for each experiment.}

@@

\subsection{Selected architecture}
The design of an neural network architecture highly depends on the problem at hand. In our case, we are trying to discriminate samples of two distributions cosmologicals maps. Hence, we assume that each pixel carries some information about the distribution. Thanks to 6 graph convolution layers (GC sizes: 16, 32, 64, 64, 64, 2) we first build the stistical evidence of each distribution for each pixel of a downsampled sphere. The last graph convolution layer is compose of 2 channels as we have 2 distributions. Then we use a statistical 'mean' layer (SL) to average the evidence for each distribution and we compute the probabilities thanks to a softmax layer (SM). We use batch normalization (BN), pooling (P) and relu non-linearity $\sigma$ for each convolutional layer. The network can be summarized as
\begin{equation}
NN = SM \circ ST_{\text{mean}} \circ L_6 \circ L_5 \circ L_4 \circ L_3 \circ L_2 \circ L_1
\end{equation}
where
\begin{itemize}
    \item layer 1: $L_1 = P  \circ \sigma \circ BN  \circ GC_{16}$,
    \item layer 2: $L_2 = P  \circ \sigma \circ BN  \circ GC_{32}$,
    \item layer 3-5: $L_3 = L_4 = L_5= P \circ  \sigma \circ BN  \circ GC_{64}$,
    \item layer 6: $L_6 = \sigma \circ BN  \circ GC_{2}$.
\end{itemize}

\todo{We use the ADAM optimizer, \cite{kingma2014adam}, with an initial learning rate of $2e-4$ that is exponentially decreased by $0.98$ a $100$ times to end at $2.7e-5$.}
